---
title: "Gapminder analysis"
author: "Gurpreet"
date: "`r Sys.Date()`"
output: 
  html_document:  
   toc: yes
   toc_float: yes
   code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the first line of my document. In this document, I will analyze how life expectancy has changed over time.

#Reading in gapminder data

```{r reading in data}
gapminder_df <- read.csv("data/gapminder_data.csv")

```

##A smaller header

### Even smaller

#### Very small

Lists:

\*List 1

\*List 2

Another type of list:

-   Hyphen list

-   Hyphen list

Numbered lists:

1.  Number one

2.  Number two

    \*Sub list

3.  Number three

Test if we need a space

-   Test list 1

-   Test list 2

We make text **bold with two asterisks** or *italic with one asterisks*

**This is bold text.**

We can also insert greek letters $\mu$ or $\Delta$

We can make [hyperlinks](https://pad.carpentries.org/Spring2024-BIOMI6300)

#Calculating summary statistics

```{r calcualting-median-pop, echo= FALSE}
med_pop <- median(gapminder_df$pop)

num_obs <- nrow(gapminder_df)
```

Over `r num_obs` observations, the median population was `r med_pop/10e6` million people.

Over `r num_obs` observations, the median population was `r round(med_pop/10e6, 3)` million people.

#Plotting life expectancy over the years

```{r line-graph}
library(ggplot2)

ggplot(data = gapminder_df, 
       mapping = aes(x = year, y = lifeExp, group = country, color = continent)) +
         geom_line() +
         scale_y_log10()
```

#Plotting population distribution

```{r histogram, message=FALSE, fig.width = 10}
ggplot(gapminder_df, aes(x = pop)) + 
  geom_histogram()
```


